---
title: "ESM 244 Lab 1"
author: "Caitie Reza"
date: "1/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(sf)
library(tmap)
```


# Read in SF data

```{r}

sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```



### Basic Wrangling

Find top 5 highest observations of trees by legal_status, make a graph

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  relocate(tree_count) %>% 
  slice_max(tree_count, n = 5)

```

Make a graph of top 5 observations by legal status

```{r}
ggplot(data= top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip()

```

### A few more data wrangling refreshers

Only wnat to keep observations (rows) for Acacia trees


```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) +
  geom_point()
  

```


## separate() and unite() functions for combiniing/separating columns


```{r}
sf_trees_sp <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")

```


Example: tidyr::unite()


```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")

```

### Make some actual maps

Use 'st_as_sf()' to convert lat and long to spatial coordinates

```{r}
blackwood_acacia_sp <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c("longitude", "latitude"))

st_crs(blackwood_acacia_sp) = 4326

ggplot(data = blackwood_acacia_sp) +
  geom_sf(color = "darkgreen")

```
### Read in SF Roads shapefile


```{r}

sf_map <- read_sf(here("data", "sf_map", "tl_2017_06075_roads.shp"))

st_transform(sf_map, 4326)

ggplot(data = sf_map) +
  geom_sf()
```

Combine blackwood acacia tree obs with SF roads

```{r}
ggplot() +
  geom_sf(data = sf_map, size = .1, color = "darkgray") +
  geom_sf(data = blackwood_acacia_sp, color = "red", size = .5)


```


### Using tmap to create interactive map

```{r}
tmap_mode("view")

tm_shape(blackwood_acacia_sp) +
  tm_dots()

```

